sWGS Normals

### 2019-09-19 Raw Data Comparison.Rmd
...

```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata')
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

@@

@@

@@

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin scaled tile values', bottom='Val glob = validation bins normalized using global z-scores')
```

@@

>>>>

```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata',verbose=T)
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

@@

In this case the z-scores are drawn from the entire cohort similarly to what I would do before training a model with all samples.

@@

tr.int.norm = as.matrix(training.tiles[,-1])
for (i in 1:ncol(tr.int.norm)) 
  tr.int.norm[,i] = BarrettsProgressionRisk:::unit.var(tr.int.norm[,i], v.means[i], v.sd[i])

tr.int.norm = as_tibble(tr.int.norm) %>% add_column('Sample'=training.tiles$Sample,.before=T)
trinorm.dist = tile.dist.measure(tr.int.norm) %>% mutate(cohort='training glob')

@@

gridExtra::grid.arrange(
  ggplot(bind_rows(trinorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(trinorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin scaled tile values', bottom='Val glob = validation bins normalized using global z-scores')
```

@@

Starts looking a little better in non-progressors.

### 2019-09-19 sWGS Normals.Rmd
...

library(Matrix)
library(glmnet)
library(ggplot2)
library(gridExtra)
library(pander)
source('lib/load_patient_metadata.R')
source('lib/data_func.R')
source('lib/fastPCF.R')

data.files = list.files('~/Data/Ellie/QDNAseq', full.names = T)

rpi = read.patient.info('~/Data/Ellie/QDNAseq/All_patient_info.xlsx')
normals = rpi$removed
path = read_excel('~/Data/Ellie/QDNAseq/All_patient_info.xlsx', 'Normals')[,c('Patient ID', 'Hospital Research ID', 'Pathology notes','Plate Index','SLX ID 1')]
colnames(path)[c(1,4,5)] = c('Patient','Plate.Index','SLX.ID')
path$SLX.ID = sub('SLX-','', gsub(' ', '', path$SLX.ID))
normals = merge(normals, path, by=c('Patient','Plate.Index','SLX.ID'))

data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, value=T)
if (length(data.file) <= 0)
  stop("Missing necessary Rdata file from Multisample script")
load(file=data.file, verbose=T)

loc.cols = c('location','chrom','start','end')
fit.data = fit.data[,c(loc.cols,normals$Samplename)]
raw.data = raw.data[,c(loc.cols,normals$Samplename)]

exclude.file = grep('qDNAseq_blacklistedRegions.txt', data.files, value=T)
if (length(data.file) <= 0)
  stop(paste("Missing necessary exclusion file from 'qDNAseq_blacklistedRegions.txt' in",data))
blacklisted.regions = read.table(exclude.file,sep="\t",header=T,stringsAsFactors=F)

if (file.exists('/tmp/normals.Rdata')) {
  load('/tmp/normals.Rdata', verbose=T)
} else {
  plist = list()
  seg.residuals = list()
  tiled.normals = NULL; tiled.normal.arms = NULL
  for (sample in normals$Samplename) {
    binned = binSWGS(raw.data[,c(loc.cols,sample)],fit.data[,c(loc.cols,sample)],blacklisted.regions,min.probes=67,gamma2=250,metrics.file=NULL)
    plist[[sample]] = binned$plots[[1]]
    seg.residuals[[sample]] = binned$resvar[[sample]]

    tiled = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),verbose=T)
    tiled.arms = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),size='arms', verbose=T)
    if (is.null(tiled.normals) | is.null(tiled.normal.arms)) {
      tiled.normals = tiled
      tiled.normal.arms = tiled.arms
    } else {
      tiled.normals = cbind(tiled.normals, tiled[,sample,drop=F])
      tiled.normal.arms = cbind(tiled.normal.arms, tiled.arms[,sample,drop=F])
    }
  }
  save(tiled.normals, tiled.normal.arms, plist, seg.residuals, file='/tmp/normals.Rdata')
}

@@

```{r, echo=F}
modeldir = '~/Data/Ellie/Analysis/5e6_arms_all'
file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(dysplasia.df,coefs,plots,labels)

>>>>

library(tidyverse)
library(BarrettsProgressionRisk)
library(glmnet)
library(gridExtra)
library(pander)
source('lib/load_patient_metadata.R')
#source('lib/data_func.R')
#source('lib/fastPCF.R')

data.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis'
#data.files = list.files(, full.names = T)

modeldir = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms_all'

patient.file = list.files(data.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(data.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

rpi = read.patient.info(patient.file, demo.file)
normals = rpi$normal

data.files = list.files(paste0(analysis.dir, '/normals'), 'segment.Rdata', recursive = T, full.names = T)

resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))

@@

```{r, echo=F}

file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(dysplasia.df,coefs,plots,labels)
